---
title: "Morning notes"
author: "Jared Petry"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

## Nested For Loops

for loop within a for loop! outermost loop corresponds to the outermost variables innermost for loop is evaluated until it is done

'i' will start at 1 while 'j' goes 1, 2, 3, 4, 5 then 'i' will go to 2 and the same thing will happen

```{r}
file_prefix <- c("temp", "ph", "salinity") 
file_suffix <- c(1,2,3,4,5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}

#---- so what if you switch i and j?
for (j in seq_along(file_prefix)) {
  for (i in seq_along(file_suffix)) {
    print(paste0(file_prefix[j], "_", file_suffix[i]))
  }
}

#nothing happens unless you switch the relative position of i and j

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[j], "_", file_suffix[i]))
  }
}

#there are some NA values becuase the loop is trying to combine the 5th element in one vector that doesn't exist 

```

```{r}
odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

#now practice making a four loop that multiplies the first element of odds with all the individual elements in even beofre moving to the next 'odd' element and repeating the same thing


```

# Writing Functions

you should consider writing a function when you have copied and pasted a block of code more than twice
- people don't always do this 
- a function name is important ... have a good name for readabililty 
- good to name functions verbs and onjects nouns 
- general setup for functions look the same for most languages 

```{r}
#----example of a function in R 

exclaim_age <- function(age) {    #---this says function_name <- function(inputs_needed_for_func)
  print(paste("I am", age, "years old!"))      #---- body of the function... what its doing 
}

exclaim_age(age = 10)  #----- now you can call the function and put in the input that you want
  
  
```

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for (i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] +quarter_splits[i + 1]
  print(half_time)
}
```

## Combining functions with if else statements (conditionals)
```{r}
animal_age <- function(animal, age) {
  if (animal == "dog") {
    print(age * 7) 
  } else if (animal == "goat") {
    print(age*4)
  }
  }
}
```

Make a function to double the user input 
```{r}
double_this <- function(my_value) {
  print(my_value *2) 
}

double_this(4) #---if you don't specify the arguments, they have to be in exact order
double_this(my_value = 4) # if you do specify arguments, you can use any order 

#---- we can just pipe a value into this function too!
this_value <- 45

this_value |> double_this()
```

Write a function that asks for a users NAME and AGE and returns an exclamation like "I'm [NAME], and I am [age] years old!"

```{r}
name_age <- function(name, age) {
  print(paste("I'm", name, "and I am", age, "years old!"))
}

name_age(name = "jared", 23)
```

Function documentation 
title 
description - what the function does 
usage - 
usually looking in the R documentation is confusing because its for expert coders 



Now lets do some more stuff

provide a "switch" within my function... where it can see which animal it is and then choose what to print based on that 
```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age * 4.7)
  }
}

convert_animal_age(animal = "goat", human_age = 23)
# --- what if you don't put in either goat  or dog? 
convert_animal_age(animal = "jackalope", human_age = 5) #---silent failure! 

# ----- add another else if statement to catch the things that are not a valid input

convert_animal_age2 <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age * 4.7)
  } else {
    print("not a valid animal for this function")
  }
}

#now try entering jackalope

convert_animal_age2(animal = "jackalope", human_age = 23)
```

Now we are going to pull info from a data frame and this changes the parameters

take data frame and pipe into filter(condition)
OR we can...
filter(.data= dataframe, condition)
```{r}
#--- use the tribble function.. so add the tidyverse library
#--- this will be different dog types and their food choice

dog_choice <- tribble(
  ~dog_name, ~food, 
  "khora", "everything", 
  "teddy", "salmon", 
  "waffle", "pancakes", 
  "banjo", "chicken"
)

#---- do some data wrangling within our function 
dog_menu <- function(name) {  #--- name is an object just existing in the function 
  my_sub <- filter(.data = dog_choice, dog_name == name) 
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub$food))
}

#--- only keeping rows from df where the input name matches a dog name

dog_menu("waffle")
  
```


## Functions meet for loops 

getting a mean of a column for a dataframe and then printing a statement about the mean 
we are basically constructing the across() function
```{r}
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i]) 
    print(paste("the mean value of", column_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)

means_over_columns(df = faithful)

#now we don't have to copy and paste every time we use this 
```

