---
title: "Morning notes"
author: "Jared Petry"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nested For Loops

for loop within a for loop! outermost loop corresponds to the outermost variables innermost for loop is evaluated until it is done

'i' will start at 1 while 'j' goes 1, 2, 3, 4, 5 then 'i' will go to 2 and the same thing will happen

```{r}
file_prefix <- c("temp", "ph", "salinity") 
file_suffix <- c(1,2,3,4,5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}

#---- so what if you switch i and j?
for (j in seq_along(file_prefix)) {
  for (i in seq_along(file_suffix)) {
    print(paste0(file_prefix[j], "_", file_suffix[i]))
  }
}

#nothing happens unless you switch the relative position of i and j

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[j], "_", file_suffix[i]))
  }
}

#there are some NA values becuase the loop is trying to combine the 5th element in one vector that doesn't exist 

```

```{r}
odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

#now practice making a four loop that multiplies the first element of odds with all the individual elements in even beofre moving to the next 'odd' element and repeating the same thing


```

# Writing Functions

you should consider writing a function when you have copied and pasted a block of code more than twice
- people don't always do this 
- a function name is important ... have a good name for readabililty 
- good to name functions verbs and onjects nouns 
- general setup for functions look the same for most languages 

```{r}
#----example of a function in R 

exclaim_age <- function(age) {    #---this says function_name <- function(inputs_needed_for_func)
  print(paste("I am", age, "years old!"))      #---- body of the function... what its doing 
}

exclaim_age(age = 10)  #----- now you can call the function and put in the input that you want
  
  
```

